Statically Typed Web Programming using Scala

### 2012-12-21 papers/ecoop2013/repairing.tex
packing and adding some words about Links

\begin{abstract}
Rich Internet Applications rely on ever more code on the client-side, typically written in JavaScript, dealing with
DOM manipulation, event handling and asynchronous calls to the server. Beyond well documented JavaScript programming
intrinsic issues, the hard wired distribution of an application business logic across clients and servers hampers its
flexibility and maintainability. This paper presents a Scala compiled embedded DSL for Web programming trying to
address these shortcomings. Close to JavaScript, it provides type-safe and expressive abstractions for DOM
manipulation, event handling and asynchronous programming. Code written using the DSL can be compiled to both server
and client sides. Built on top of the Lightweight Modular Staging framework, this DSL uses both type-level information
and staging to  introduce high-level abstractions that translate to efficient JavaScript and Scala code using native
APIs.
\end{abstract}

{\bf Keywords:} Web, programming languages, embedded DSL, type-level programming, staging, Scala

\section{Introduction}

The Web is an appealing platform on which to write applications (\aka{} Rich Internet Applications, RIA) because it
makes them easy to deploy on clients and allows large scale innovative collaborative
experiences~\cite{Farrell07_RIA,Mikkonen08_SpaghettiJs}. RIAs are characterized by partial updates of the user
interface (as opposed to refreshing the whole page with the classic hyperlink navigation), so a large part of the
client-side code listens to user interface events (such as mouse clicks), triggers the appropriate action
asynchronously on the server using AJAX~\cite{Garrett05_Ajax} and updates the application’s state and the page content
with a new DOM fragment computed from the response data~\cite{Farrell07_RIA,Busch09_StateOfArtRIA}.

@@

Opa\footnote{\href{http://opalang.org/}{http://opalang.org/}} is a language and a full stack designed for Web
programming. It features static typing with global type inference, modules and pattern matching. The language
provides a syntactic sugar for writing HTML fragments in the code. The server and client parts of an application can
be written in the Opa language, the system can decide whether a code fragment will run on server-side, client-side or
both. However, the Opa language is not extensible, so there is no way for developers to customize the compilation
process of their abstractions on client and server sides.

>>>>

\begin{abstract}
\vspace*{-0.7cm}
Rich Internet Applications rely on ever more code on the client-side, typically written in JavaScript, dealing with
DOM manipulation, event handling and asynchronous calls to the server. Beyond well documented JavaScript programming
intrinsic issues, the hard wired distribution of an application business logic across clients and servers hampers its
flexibility and maintainability. This paper presents a Scala compiled embedded DSL for Web programming trying to
address these shortcomings. Close to JavaScript, it provides type-safe and expressive abstractions for DOM
manipulation, event handling and asynchronous programming. Code written using the DSL can be compiled to both server
and client sides. Built on top of the Lightweight Modular Staging framework, this DSL uses both type-level information
and staging to  introduce high-level abstractions that translate to efficient JavaScript and Scala code using native
APIs.
\end{abstract}
\vspace*{-0.3cm}
{\bf Keywords:} Web, programming languages, embedded DSL, type-level programming, staging, Scala

\section{Introduction}
\vspace*{-0.2cm}
The Web is an appealing platform on which to write applications (\aka{} Rich Internet Applications, RIA) because it
makes them easy to deploy on clients and allows large scale innovative collaborative
experiences~\cite{Farrell07_RIA,Mikkonen08_SpaghettiJs}. RIAs are characterized by partial updates of the user
interface (as opposed to refreshing the whole page with the classic hyperlink navigation), so a large part of the
client-side code listens to user interface events (such as mouse clicks), triggers the appropriate action
asynchronously on the server using AJAX~\cite{Garrett05_Ajax} and updates the application’s state and the page content
with a new DOM fragment computed from the response data~\cite{Farrell07_RIA,Busch09_StateOfArtRIA}.

@@

Opa\footnote{\href{http://opalang.org/}{http://opalang.org/}} and Links~\cite{Cooper07_Links} are languages designed
for Web programming. They feature static typing with global type inference, modules and pattern matching. These
languages provide a syntactic sugar for writing HTML fragments directly in the code. The server and client parts of
an application are written in a single language and the system can decide whether a code fragment will run on
server-side, client-side or both. However, these languages are not extensible, so there is no way for developers to
customize the compilation process of their abstractions on client and server sides. Furthermore, hiding the code
execution distribution gives programmers less control on communication error handling~\cite{Guerraoui99_Distributed}.
